State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-16 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
more paper edits

@@

@@

The width parameter of the Gaussian basis functions can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. Alternatively, the upper spatial frequency content of the neural field modelled by the basis function decomposition can be determined, given the basis function width. Since Shannon's sampling theorem is derived for perfect point sensors, this is particularly useful when considering spatially distributed basis functions. To demonstrate this, we begin with the two-dimensional Gaussian basis function defined as
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
with the corresponding Fourier transform
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{cy}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent the field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

Note that for spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Thus, the knowledge of the distance between basis functions can be used to determine the total number of basis functions required to represent a known spatial region.

@@

The sensors were spaced in a regular square 14 $\times$ 14 grid, the sensor width $\sigma^2_m$ was $0.81$, which equals $1.5$~mm full width at half maximum, with spacing $\Delta_y$ equal to $1.5$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. 
\begin{figure}
   	\begin{center}
    		\subfigure[][]{\includegraphics{./Graph/ExperimentFigureEEGPy.pdf}} \\
   		\subfigure[][]{\includegraphics{./Graph/ExperimentFigurePSDPy.pdf}} \\
   	\end{center}
   	\caption{\textbf{a.} Example of data generated by the full neural field model from the first 5 sensors. \textbf{b.} Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 
\label{fig:experimental design}
\end{figure}
\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT} shows the spatial frequency of the neural field. The cut-off frequency, $\nu_c$, was taken to be 0.24~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of 2.05~mm. This confirms the separation, $\Delta_{y}$, of 1.5~mm was sufficient to prevent problems associated with spatial aliasing.
\begin{figure}
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics{./Graph/FFTEstimatedField.pdf}}\\
\caption{Power in dB as a function of spatial frequency of the neural field. \textbf{a.} The average (over time) power of the spatial frequency of the neural field. \textbf{b.} The average (over time) power of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}
\label{fig:FFTTrueEstimate}
\end{figure}
The spatial frequency from the observations is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency, $\nu_{cy}$, was taken to be $0.19$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter, $\rho_{\phi}=1$, the minimum distance between basis functions was calculated to be $2.63$ to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions, $\Delta_{\phi}$, was set to $2.5$, which is sufficient to capture the significant dynamics according to our criteria.

Given the basis function separation, the basis function width was set to $\sigma_{\phi}^2=2.5$~mm. This facilitated a smooth representation of the field by the basis function decomposition given the basis function separation and provided low-pass spatial filtering of the field. From equation~\ref{eq:WidthCutOffRelationship}, this allowed the basis function decomposition to represent spatial frequencies up to approximately $0.12$~Hz. The spatial frequency response of the basis functions can be seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
\begin{figure} 
	\begin{center}
		\subfigure[][]{\label{fig:ObservationFreqAnal}\includegraphics{./Graph/ObservationFrequencyResponse.pdf}}\\	
		\subfigure[][]{\label{fig:SensorFunctionFreqResponse}\includegraphics{./Graph/SensorChar.pdf}}
		\subfigure[][]{\label{fig:BasisFunctionFreqResponse}\includegraphics{./Graph/BasisChar.pdf}}
	\end{center}
	\caption{Spatial frequency analysis. \textbf{a.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field. The symmetry of the sensors and field basis functions allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b.} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} The spatial frequency response of a 1-dimensional neural field basis function.} 
\end{figure}

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the parameter and state estimation, a Monte Carlo approach was employed where $100$ realisations of the data were generated. Each realisation consisted of $500$~ms of data (sampled at $1$~kHz) and the estimation was applied to the final $400$~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. The distribution of the estimated parameters is shown in Figure~\ref{fig:Parameters}.

The estimates for $\xi$, where $\xi=1-\zeta T_s $ and $\zeta$ is the  reciprocal of the synaptic time constant, are consistently over the true value (with a mean bias of ?? \%). The distribution of the  estimates of the connectivity kernel parameters ($\theta_0$, $\theta_1$, $\theta_2$) are centred about the true values, with standard deviation \dean{??}.  \dean{ for discussion: indicating good results}. The ratios of the estimates ($\hat\theta_0/\hat\theta_1$ and $\hat\theta_0/\hat\theta_2$) are shown in Figure~\ref{fig:ParametersRatio}. The distribution of estimates for $\hat\theta_0/\hat\theta_1$ are relativity tighter, though indicate a bias... This indicates the balance of excitation and inhibition can be estimated more accurately than the individual parameters.

@@

Figure~\ref{fig:ParametersConvergence} shows the rate of convergence for parameter estimates.
\begin{figure}
    \centering
    \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
    \subfigure[][]{\includegraphics{./Graph/theta1.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameter estimates over 100
realisations. \textbf{a.} The distribution of $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the synaptic time constant (true parameter: $\xi=0.9$). \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates (true parameter: $\theta_0 = 10$). \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates (true parameter: $\theta_1 = -8$). \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates (true parameter $\theta_3 = 0.5$). The dotted line shows the location of the true parameter in each case.}
\label{fig:Parameters}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). The dotted line shows the location of the true ratio in each case.}
\label{fig:ParametersRatio}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/KernelEstimate.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Results for connectivity kernel estimate. In all the subplots the true kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dashed line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain.}
\label{fig:KernelEstimates}
\end{figure}
\begin{figure}
        \centering
\includegraphics{./Graph/convergence.pdf}
\caption{The mean of the absolute error between true and estimated parameters averaged over 100 runs. All parameters converge after six iterations. The mean difference between errors in consecutive iterations falls below $10^{-4}$ after 6 iterations. }
\label{fig:ParametersConvergence}
\end{figure}

@@

This estimation framework also has theoretical implications by establishing a framework that has the potential to allow testing of hypotheses that have been generated in theoretical studies and to validate neural field models. For example, it has been hypothesised that so-called ``bump solutions'' are a possible mechanism for short term memory formation~\cite{Coombes2005}. Using the proposed model-based framework, we can reconstruct a neural field from data, and estimate parameters and check if they correspond the the theoretically derived parameters where these bump solutions can exist. In addition, the parameters and the state-space can be explored during other cognitive tasks and states of arousal such as during sleep and anaesthesia.

\section{Conclusion}

This paper presents a novel framework for state and parameter estimation of spatio-temporal neural fields. A basis function decomposition reduces the infinite dimensional neural field (continuous space) to a finite dimensional state-space model. The level of detail of the estimated field is related to the basis functions decomposition, providing a flexible framework where a trade-off can be made between physiological detail and computational complexity. The effect of the basis function decomposition will restrict the frequency content of the estimated field. This is evident in when comparing Figures~\ref{fig:FieldFFT} and~\ref{fig:EstimatedFieldFFT} where the estimated field is spatially band-limited. The spatial frequency content of the field is governed by disturbance covariance and the shape of the connectivity kernel. Using the state-space model, the unscented Rauch Tung Striebel smoother was applied for state estimation. Using the sequence of state estimates, a least squares algorithm was implemented for parameter estimation. We have demonstrated that it is theoretically possible to estimate intracortical connectivity and the synaptic time constant.

The frequency analysis and model selection tools presented in this paper specify the sensor and field basis function arrangement required to form a state-space model that is capable of capturing the dominant cortical dynamics. In addition to providing conditions to ensure the reduced model can capture the significant spectral properties of the field, these new tools can be used to better design electrophysiological recording techniques avoid spatial aliasing. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009}.

In this paper,\parham{In this paper, assumption was made that the chosen kernel basis functions can exactly represent the actual kernel. In future work, we plan to extent this work to relax this assumption. We hope to achieve this by using the actual kernel that cannot exactly be presented by the chosen finite number of basis functions. Visakan: width estimation is totally different process} the widths of the connectivity kernel basis functions are assumed to be known. In future work, we plan to extend this work to estimate the connectivity structure with unknown kernel basis function widths. We hope to achieve this by using a sufficiently large set of connectivity basis functions within the physiological plausible limits, and allow the estimation procedure take care of the weights. In addition, we plan to show in future work how the assumption of homogeneity can be relaxed, where heterogeneity can be modelled by careful placement of field basis functions. For example, if cortical regions have a greater density of neural populations, then more basis functions can be used in these areas \parham{Visakan: heteroginity is about the process not presentation, and it should be addressed by using a kernel which is different at different spatial locations}. In this scenario, model selection and basis function positioning can be achieved using instantaneous frequency analysis of the the neural field or alternative model selection tools based on information criteria. Future work should also be targeted to include a more realistic synaptic response kernel, with finite rise and decay times and finite propagation times into the neural field equations.

@@

\section{Discrete Time Model}\label{Time Discretization} To form the IDE neural field model a time discretisation must be perform. We used a one-step Euler method where Eq.~\ref{FinalFormContinuous} can be approximated by 

>>>>

	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t

@@

and 

@@

The width parameter of the Gaussian basis functions can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. Alternatively, the upper spatial frequency content of the neural field modelled by the basis function decomposition can be determined, given the basis function width. Since Shannon's sampling theorem is derived for perfect point sensors, this is particularly useful when considering spatially distributed basis functions. To demonstrate this, we begin with the two-dimensional Gaussian basis function defined as
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
with the corresponding Fourier transform
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{cy}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent the neural field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

Note that for a spatially homogeneous isotropic field, the basis functions can be placed on a regular grid. Thus, the knowledge of the distance between basis functions can be used to determine the total number of basis functions required to represent a known spatial region.

@@

The sensors were spaced in a regular square 14 $\times$ 14 grid, the sensor width $\sigma^2_m$ was $0.81$, which provides a width of $1.5$~mm at half the maximum amplitude of the basis function, with spacing $\Delta_y=1.5$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. 
\begin{figure}
   	\begin{center}
    		\subfigure[][]{\includegraphics{./Graph/ExperimentFigureEEGPy.pdf}} \\
   		\subfigure[][]{\includegraphics{./Graph/ExperimentFigurePSDPy.pdf}} \\
   	\end{center}
   	\caption{\textbf{a.} Example of data generated by the full neural field model from the first 5 sensors. \textbf{b.} Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 
\label{fig:experimental design}
\end{figure}
\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT} shows the spatial frequency of the neural field. The cut-off frequency, $\nu_c$, was taken to be 0.24~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of 2.05~mm. This confirms the separation, $\Delta_{y}$, of 1.5~mm was sufficient to prevent problems associated with spatial aliasing.
\begin{figure}
\centering
\subfigure[][]{\label{fig:FieldFFT}\includegraphics{./Graph/FFTTrueField.pdf}}
\subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics{./Graph/FFTEstimatedField.pdf}}\\
\caption{Power in dB as a function of spatial frequency of the neural field. \textbf{a.} The average (over time) power of the spatial frequency of the neural field. \textbf{b.} The average (over time) power of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}
\label{fig:FFTTrueEstimate}
\end{figure}
The spatial frequency from the observations is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency, $\nu_{cy}$, was taken to be $0.19$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter, $\rho_{\phi}=1$, the minimum distance between basis functions was calculated to be $2.63$ to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions, $\Delta_{\phi}$, was set to $2.5$, which is sufficient to capture the significant dynamics according to our criteria.

Given the basis function separation, the basis function width was set to $\sigma_{\phi}^2=2.5$~mm. This facilitated a smooth representation of the field by the basis function decomposition given the basis function separation and provided low-pass spatial filtering of the field. From equation~\ref{eq:WidthCutOffRelationship}, this allowed the basis function decomposition to represent spatial frequencies up to approximately $0.12$~Hz. The spatial frequency response of the basis functions can be seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
\begin{figure} 
	\begin{center}
		\subfigure[][]{\label{fig:ObservationFreqAnal}\includegraphics{./Graph/ObservationFrequencyResponse.pdf}}\\	
		\subfigure[][]{\label{fig:SensorFunctionFreqResponse}\includegraphics{./Graph/SensorChar.pdf}}
		\subfigure[][]{\label{fig:BasisFunctionFreqResponse}\includegraphics{./Graph/BasisChar.pdf}}
	\end{center}
	\caption{Spatial frequency analysis. \textbf{a.} Two-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field. The symmetry of the sensors and field basis functions allow for the 1-dimensional representation of the spectral characteristics. \textbf{b.} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} The spatial frequency response of a 1-dimensional neural field basis function. The dashed line shows the cut-off frequency (-3 dB point).} 
\end{figure}

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the parameter and state estimation, a Monte Carlo approach was employed where $100$ realisations of the data were generated. Each realisation consisted of $500$~ms of data (sampled at $1$~kHz) and the estimation was applied to the final $400$~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Six iterations of the estimation algorithm were required for the parameters to converge.

The distribution of the estimated parameters is shown in Figure~\ref{fig:Parameters}. The estimates for $\xi$, where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant, are consistently over the true value (with a mean bias of \dean{??~\%}). The estimates of the connectivity kernel parameters ($\theta_0$, $\theta_1$, $\theta_2$) form distributions that are approximately Gaussian, which centred about the true values, with standard deviation \dean{??}.  \dean{ for discussion: indicating good results}. The ratios of the estimates ($\hat\theta_0/\hat\theta_1$ and $\hat\theta_0/\hat\theta_2$) are shown in Figure~\ref{fig:ParametersRatio}. The distribution of estimates for $\hat\theta_0/\hat\theta_1$ are relativity tighter, though indicate a bias... This indicates the balance of excitation and inhibition can be estimated more accurately than the individual parameters.

@@

Figure~\ref{fig:ParametersConvergence} shows the rate of convergence for parameter estimates.
\begin{figure}
    \centering
    \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
    \subfigure[][]{\includegraphics{./Graph/theta1.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameter estimates over 100
realisations. The dotted line shows the location of the true parameter in each case. \textbf{a.} The distribution of $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the synaptic time constant (true parameter: $\xi=0.9$). \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates (true parameter: $\theta_0 = 10$). \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates (true parameter: $\theta_1 = -8$). \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates (true parameter $\theta_3 = 0.5$).}
\label{fig:Parameters}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). The dotted line shows the location of the true ratio in each case.}
\label{fig:ParametersRatio}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/KernelEstimate.pdf}}
\subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Results for connectivity kernel estimate. In all the subplots the true kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the 95~\% confidence interval is shown by the shaded grey region. \textbf{a.} Comparison of kernel estimates to true kernel in the spatial domain. \textbf{b.} Comparison of the true kernel to the estimated kernel in the frequency domain.}
\label{fig:KernelEstimates}
\end{figure}
\begin{figure}
        \centering
\includegraphics{./Graph/convergence.pdf}
\caption{The mean of the absolute error between true and estimated parameters averaged over 100 runs. All parameters converge after six iterations. The mean difference between errors in consecutive iterations falls below $10^{-4}$ after 6 iterations. }
\label{fig:ParametersConvergence}
\end{figure}

@@

This estimation framework also has theoretical implications by establishing a framework that has the potential to allow testing of hypotheses that have been generated in theoretical studies and to validate neural field models. For example, it has been hypothesised that so-called ``bump solutions'' are a possible mechanism for short term memory formation~\cite{Coombes2005}. Using the proposed model-based framework, we can reconstruct a neural field from data, and estimate parameters and check if they correspond to the theoretically derived parameters where these bump solutions can exist. In addition, the parameters and the state-space can be explored during other cognitive tasks and states of arousal such as during sleep and anaesthesia.

\section{Conclusion}
This paper presents a novel framework for state and parameter estimation of spatio-temporal neural fields. A basis function decomposition reduces the infinite dimensional neural field (continuous space) to a finite dimensional state-space model. The level of detail of the estimated field is related to the basis functions decomposition, providing a flexible framework where a trade-off can be made between physiological detail and computational complexity. The effect of the basis function decomposition will restrict the frequency content of the estimated field. This is evident when comparing Figures~\ref{fig:FieldFFT} and~\ref{fig:EstimatedFieldFFT} where the estimated field is spatially band-limited. The spatial frequency content of the field is governed by disturbance covariance and the shape of the connectivity kernel. Using the state-space model, the unscented Rauch Tung Striebel smoother was applied for state estimation. Using the sequence of state estimates, a least squares algorithm was implemented for parameter estimation. We have demonstrated that it is theoretically possible to estimate intracortical connectivity and the synaptic time constant.

The frequency analysis and model selection tools presented in this paper specify the sensor and field basis function arrangement required to form a state-space model that is capable of capturing the dominant cortical dynamics. In addition to providing conditions to ensure the reduced model can capture the significant spectral properties of the field, these new tools can be used to better design electrophysiological recording techniques to avoid spatial aliasing. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009}.

In this paper,\parham{In this paper, assumption was made that the chosen kernel basis functions can exactly represent the actual kernel. In future work, we plan to extent this work to relax this assumption. We hope to achieve this by using the actual kernel that cannot exactly be presented by the chosen finite number of basis functions. Visakan: width estimation is totally different process} the widths of the connectivity kernel basis functions are assumed to be known. In future work, we plan to extend this work to estimate the connectivity structure with unknown kernel basis function widths. We hope to achieve this by using a sufficiently large set of connectivity basis functions within the physiological plausible limits, and allow the estimation procedure to take care of the weights. In addition, we plan to show in future work how the assumption of homogeneity can be relaxed, where heterogeneity can be modelled by careful placement of field basis functions. For example, if cortical regions have a greater density of neural populations, then more basis functions can be used in these areas \parham{Visakan: heteroginity is about the process not presentation, and it should be addressed by using a kernel which is different at different spatial locations}. In this scenario, model selection and basis function positioning can be achieved using instantaneous frequency analysis of the the neural field or alternative model selection tools based on information criteria. Future work should also be targeted to include a more realistic synaptic response kernel, with finite rise and decay times and finite propagation times into the neural field equations.

@@

\section{Discrete Time Model}\label{Time Discretization} To form the IDE neural field model a time discretisation must be performed. We used a one-step Euler method where equation~\ref{FinalFormContinuous} can be approximated by 
